Spring haul-out behavior of seals in the Bering and Chukchi seas

### 2023-03-01 analysis/paper/reviewer_response.md

> *Perhaps you mean to offer a topic sentence, stating "How
> ice-assoicated seals use sea ice as a haulout platform varies betweens
> species".*

We have adopted this suggestion

> There is so much to state about ice-associated seals in the Pacific
> Arctic that this reference to Antarctic work seems out of place here.

We have moved this reference to later in the paragraph so the first
references are about Arctic work. While the example is from the
Antarctic, it is an important, early example of analysis using
bio-logger data to estimate haul-out behavior in seals.

@@

This sentence now reads:

"Ship-based capture events occurred during spring near the southern ice
edge in the Bering Sea between 2005 and 2018."

> Please provide the geographic context of your tag deployments and data
> collection using a plot similar to that used by Jay et al. (2012:
> 10.3354/meps10057 figure 2). Please consider using a tabulation that
> enables readers to understand the geographic context of the
> deployments similar to that used by Jay et al. 2012, table 1.

We feel the spatial distribution of the actual transmitted data --
compared to the deployment location -- is most informative and relevant
for the audience's understanding of how our data are distributed
spatially and how well data represents the range of each species. We
provide this visually in Figure 3 and adding an additional layer to that
figure would to show deployments would be an additional complication. To
ensure the audience as a clear understanding of our study area, we have
added an additional figure with a map of the Bering and Chukchi sea
region that indicates generally where deployments were initiated.

@@

We've edited this sentence and removed the word 'key'

> This is a procedural detail. Other readers may replicate your method
> without needing to do this. Please omit this detail.

> Readers can replicate your methods without the details on converting
> the NARR wind vectors into wind speed.

We're not sure why the reviewer would recommend omitting a 'procedural
detail' as part of our methods. We feel it is important to clearly
communicate our procedure for processing the large NARR data set and
that we used the native grid size of 32km. Regarding the calculation of
wind vectors, we include this detail because many readers, like us
initially, may not be well versed in how wind vectors components are
converted to wind speed.

@@

> Understanding pinniped haulout behavior is essential to designing and
> interpreting population estimation aerial surveys. Understanding how
> they use sea ice during a time when sea ice availability is rapidly
> changing is also important, but is ancillary to the paper's focus.

We have removed the term 'major focus' and, as the reviewer suggests,
simply state that this is the goal of the study. We have also added
mention of the statutory requirements to paragraph 2 of the Introduction
section and expanded the text to acknowledge the importance of
understanding how seals use sea ice during this period of change but,
ultimately, it's ancillary to the focus of this paper.

As for providing the reader focus on why the Federal government is doing
this study: this is a collaborative study across multiple organizations
and the analysis/purpose is not limited to the Federal government. We've
incorporated the reviewers suggests to expand discussion of the
statutory requirements of the MMPA and ESA but don't feel we need to
single out the motivations for the Federal government.

We have also added a figure depicting the estimated species range for bearded,
ribbon, and spotted seals along with general representation of recent aerial
survey effort. This should better communicate to the reader how the spatial
distribution of our haul-out behavior overlaps with the species range and
survey effort.

@@

> Seals can just as happily be on foraging trips within the pack and
> marginal sea ice as in the open water, depending on the species.
> Suggesting that foraging trips are associated with ice-free conditions
> is misleading.

We have removed 'e.g., foraging trips' from this sentence so as not to
mislead the reader.

> This is your primary reason for not including sea ice geography in
> your analysis: you simply lacked geolocation of much of your haulout
> behavior data.
>
> Please help the reader understand how coarse your tracking data was
> for the flipper tagged animals and what proportion of each species was
> tracked with these flipper tags. Please also help the reader
> understand how the tracking data relates geographically to regions
> that may be subject to aerial survey efforts (i.e. are your tracking
> data collected from across the regions that may be surveyed during
> range-wide spring aerial surveys?).

We have re-organized this paragraph to, hopefully, better communicate our
choice not to explore sea-ice related covariates. As the reviewer suggests,
we have limited geolocation data for at-sea periods of our haul-out behavior
data -- we know the seal was in the water but don't have any estimates of
location when only a flipper tag is attached. By leading with this statement,
we hope that communicates our primary rationale. But, the other rationales
are also important and we still include them. We have also moved this section
to the 'Discussion' in response to other reviewer comments.

We have provided some statistics regarding the percentage of the
percent-dry timeline data that is not associated with a corresponding location
estimate because the seal was in the water and no estimates were available from
flipper mounted SPOT tags.

We have also added a figure depicting the estimated species range for bearded,
ribbon, and spotted seals along with general representation of recent aerial
survey effort. This should better communicate to the reader how the spatial
distribution of our haul-out behavior overlaps with the species range and
survey effort.

### 2023-03-01 analysis/paper/London_HauloutBeahvior.Rmd
updated figures and response to reviewer

How ice-assoicated seals use sea ice as a haul-out platform varies between species.
Ribbon seals (_Histriophoca fasciata_) haul out of the water almost exclusively
on sea ice and are mostly pelagic outside the spring pupping, breeding, and
molting season [@boveng2018a]. While spotted (_Phoca largha_) and bearded
(_Erignathus barbatus_) seals rest on coastal features, they strongly prefer sea
ice as a haul-out platform during the spring and early summer [@frost2018].
Ringed seals (_Phoca hispida_) --- not included in this study --- haul out on
sea-ice but also within snow lairs during winter and spring.

@@

(ref:deployMapCaption) Initial deployment areas for each seal species (bearded, ribbon, and spotted seals) in the study between 2005 and 2020. Deployments were initiated across a range of months but only data from 1 March to 15 July were included in the analysis. See Figure \@ref(fig:dataMap) for the spatial distribution of observed data used in the study.

@@

deploy_sf <- deploy_details %>% 
  sf::st_as_sf(coords = c('longitude','latitude'), crs = 4326) %>% 
  sf::st_transform(3571) %>% 
  dplyr::group_by(species,location, region) %>% 
  dplyr::group_modify(~ concaveman::concaveman(., concavity = 4),.keep=TRUE) %>% 
  sf::st_sf(crs = 3571) %>% 
  sf::st_buffer(40*1000)

@@

p <- ggplot() +
  layer_spatial(deploy_sf, aes(fill=species),
                linewidth=0) +
  annotation_spatial(world,
                     linewidth=0,
                     fill = "grey80") +
  facet_grid(cols = vars(species), labeller = species_labels,
             switch="x") +
  scale_fill_manual(values = met.brewer("Isfahan2",n=3,direction=-1,override.order = TRUE),
                    aesthetics = "fill", 
                    name = "species") + 
  ggtitle("Areas of initial deployment location",
          subtitle = 
                 paste0("Bio-loggers were initially deployed on ",
                "<span style = 'color:",species_colors['Bearded seal'],";'>bearded, </span> ", 
                "<span style = 'color:",species_colors['Ribbon seal'],";'>ribbon, </span> ",  
                "and ", 
                "<span style = 'color:",species_colors['Spotted seal'],";'>spotted seals</span> ", 
                "across multiple regions within the Bering and Chukchi seas.")) +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.placement = "outside",
        strip.text.x = element_markdown(face="bold"),
        plot.subtitle = element_textbox_simple(margin = ggplot2::margin(b=12)))

>>>>

How ice-associated seals use sea ice as a haul-out platform varies between species.
Ribbon seals (_Histriophoca fasciata_) haul out of the water almost exclusively
on sea ice and are mostly pelagic outside the spring pupping, breeding, and
molting season [@boveng2018a]. While spotted (_Phoca largha_) and bearded
(_Erignathus barbatus_) seals rest on coastal features, they strongly prefer sea
ice as a haul-out platform during the spring and early summer [@frost2018].
Ringed seals (_Phoca hispida_) --- not included in this study --- haul out on
sea-ice but also within snow lairs during winter and spring.

@@

(ref:deployMapCaption) Initial deployment areas for each seal species (bearded, ribbon, and spotted seals) in the study between 2005 and 2020. Colored regions shown are minimum concave polygons buffered by 60 km for enhanced visibility. See Supplemental Material S1 for specific details. Deployments were initiated across a range of months but only data from 1 March to 15 July were included in the analysis. See Figure \@ref(fig:dataMap) for the spatial distribution of observed data used in the study.

@@

deploy_sf <- deploy_details %>% 
  sf::st_as_sf(coords = c('longitude','latitude'), crs = 4326) %>% 
  sf::st_transform(3571) %>% 
  dplyr::group_by(species,location, region) %>% 
  dplyr::group_modify(~ concaveman::concaveman(., concavity = 4),.keep=TRUE) %>% 
  sf::st_sf(crs = 3571) %>% 
  sf::st_buffer(60*1000)

@@

p <- ggplot() +
  layer_spatial(deploy_sf, aes(fill=species),
                linewidth=0) +
  annotation_spatial(world,
                     linewidth=0,
                     fill = "grey80") +
  facet_grid(cols = vars(species), labeller = species_labels,
             switch="x") +
  scale_fill_manual(values = met.brewer("Isfahan2",n=3,direction=-1,override.order = TRUE),
                    aesthetics = "fill", 
                    name = "species") + 
  ggtitle("Areas of initial deployment location",
          subtitle = 
                 paste0("Bio-loggers were initially deployed on ",
                "<span style = 'color:",species_colors['Bearded seal'],";'>**bearded**, </span> ", 
                "<span style = 'color:",species_colors['Ribbon seal'],";'>**ribbon**, </span> ",  
                "and ", 
                "<span style = 'color:",species_colors['Spotted seal'],";'>**spotted**</span> seals ", 
                "across multiple regions within the Bering, Chukchi, and western Beaufort seas.")) +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.placement = "outside",
        strip.text.x = element_markdown(face="bold"),
        plot.subtitle = element_textbox_simple(margin = ggplot2::margin(b=12)))
